### 2023-05-02 review_learning_dashboards_paper/topics_keywords_app/app.rmd
Update paper.rmd

---
title: "Explore the links between your own keywords and topics"
author: "Maria Osipenko"
date: ""
output: html_document
runtime: shiny
---

<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(quanteda)
library(tidyverse)
library(keyATM)
library(igraph)
```

Try it yourself! Type in, for example: game, play, excit

```{r}
load("works_search_filter_all.RData")
works_search<-works_search[!is.na(works_search$ab),]

#mydfm: combine titles and abstracts
texts<-works_search[,c("id","display_name","ab")] 
texts$ab<-gsub("Abstract","",texts$ab)
texts$ab<-paste(texts$display_name,texts$ab)
texts<-texts[!duplicated(texts$id),]
rm_short_tokens<-function(tk,min.len=3){
  return(tk[nchar(tk)>=min.len])
}
make_dfm<-function(texts){
  #stopwords
  mystopwords<-c("may","can","sub","now","etc", "the","via","let", "e.g", "also")
  #tokenize
  tk<-tokenizers::tokenize_word_stems(texts, stopwords = c(stopwords::stopwords("en"),mystopwords))
  tk<-lapply(tk,rm_short_tokens)
  tk<-tokens(as.tokens(tk), remove_punct=TRUE, remove_numbers = TRUE, remove_symbols = TRUE,remove_url = TRUE, remove_separators = TRUE, split_hyphens=T)
  tk<-lapply(tk,gsub,pattern="quizz",replacement="quiz"); tk<-lapply(tk,gsub,pattern="quizizz",replacement="quiz")
  tk<-lapply(tk,gsub,pattern="customis",replacement="customiz"); 
  tk<-lapply(tk,gsub,pattern="practis",replacement="practic")
  tk2<-tokens_ngrams(as.tokens(tk),n=1:2,skip=4)
  #dfm
  mydfm<-dfm(tk2, verbose = FALSE,tolower=TRUE)
  #keep only terms with more than two chars
  mydfm<-dfm_select(mydfm,pattern=c(stopwords::stopwords("en"),mystopwords),valuetype="fixed",selection="remove")
  #keep only words with frequency >=10
  mydfm<-dfm_trim(mydfm, min_docfreq = 0.001, max_docfreq=0.5,docfreq_type="prop")
  #take relative frequencies as entries
  return(mydfm)
}
mydfm<-make_dfm(texts$ab)

# choose keywords
keywords<-as.list(1:6)
names(keywords)<-c("engagement","exploration","explanation","elaboration","evaluation","evolution")
keywords$engagement<-c("engag","motiv","emot","promot") 
keywords$exploration<-c("explor", "experi", "interact", "activ")
keywords$explanation<-c("explain", "instruct","knowledg", "theori") 
keywords$elaboration<-c("problem","solv", "practic", "applic")
keywords$evaluation<-c("assess", "evalu","test","feedback") 
keywords$evolution<-c("adapt", "individu",  "improv", "behavior") 
keywords_df<-data.frame(keywords)

#load the model
load("mod.RData")
K<-nrow(mod$phi)
topwords<-top_words(mod, n = 20)
colnames(topwords)<-c(colnames(topwords)[1:6], "7_research_dir", "8_tech_accept", "9_distanc_educ")
topicprop<-apply(mod$theta,2,mean) #overall topic proportion
cols <- c("gray66", "gray55", "gray44", hcl.colors(6, palette = "viridis", alpha = 0.8)[1:5],"indianred2")
tdocs<-top_docs(mod, n = 100)
topic_prop<-mod$theta
topic_prop<-data.frame(cbind(topic_prop,year=works_search$publication_year))
colnames(topic_prop)<-c(names(keywords),"research_dir","tech_accept","distanc_educ","year")

```

```{r}
library(shiny)
shinyApp(

  ui = fluidPage(
    textInput("kw", "Keywords (separated by comma):"),
    actionButton("bkw","show the links",style="color: #fff; background-color: #337ab7; border-color: #2e6da4"),
    textOutput("kwwarn"),
    plotOutput("kwPlot",width = "100%")
  ),

  server = function(input, output) {

    mykeywords<-reactiveVal()
    counts<-reactiveVal()
    omit_text<-reactiveVal()
    netk<-reactiveVal()
    papers<-reactiveVal()
    pcols<-reactiveVal()

    observeEvent(input$bkw,{
      #keywords
      kw<-gsub("[[:space:]]", "", unlist(strsplit(input$kw,"\\,")))
      mydfm_r<-mydfm[unique(as.numeric(as.matrix(tdocs))),]
      ind0<-apply(mydfm_r,2,sum)!=0
      mydfm_r<-(mydfm_r[,ind0])
      mykeywords(kw[kw%in%colnames(mydfm_r)])
      l<-length(mykeywords())

      #counts
       if(l>0){
        coun<-matrix(0,l+K,l+K)
          for(i in 1:l){
            for(j in 1:K){
             coun[i,j+l]<-sum(mydfm[as.numeric(tdocs[,j]),mykeywords()[i]])
            }
          }
        counts(coun)

      #netk

        nett <- graph_from_adjacency_matrix(as.matrix(counts()>0))
        #colors
        kcols<-hcl.colors(max(l,2), palette = "green-orange", alpha = 0.8)[1:l]

        #vetrices size, color
        V(nett)$size<-c(rep(0.1,l),apply(topic_prop,2,mean)[1:K])*300
        V(nett)$color<-c(kcols,rev(cols)[1:K])
        V(nett)$label <- c(mykeywords(),1:K)
        V(nett)$shape <- c(rep("circle",l),rep("circle",K))
        V(nett)$label.color <- rep("black",K+l)

        #edges width
        E(nett)$width<-c(t(counts())[t(counts())>0]*0.25)
        E(nett)$arrow.mode<-"-"

        netk(nett)

        #top papers
        counts_total<-apply(counts(),2,sum)[-(1:length(mykeywords()))] # total counts of keywords per topic
        tt<-order(counts_total, decreasing = T) #sort decreasing
        howmany_topics<-5
        tt<-tt[1:min(length(tt[counts_total>0]),howmany_topics)] #take at most howmany_topics with the keywords

        #counts per paper
        pcounts<-matrix(0,nrow(tdocs),K)
        for(i in 1:nrow(tdocs)){
          for(j in 1:K){
            pcounts[i,j]<-sum(mydfm[as.numeric(tdocs[i,j]),mykeywords()])
          }
        }
        howmany_papers<-3
        ppapers<-NULL
        for(i in 1:length(tt)){
          ppapers<-c(ppapers,works_search$display_name[tdocs[order(pcounts[,i], decreasing=T),tt[i]][1:howmany_papers]])
        }

        papers(paste(rep(tt, each=howmany_papers), ppapers))

        #colors
        pcols(rep(rev(c("gray66", "gray55", "gray44", 
                         hcl.colors(6, palette = "viridis", alpha = 1)[1:5],"indianred2"))[tt],each=howmany_papers))
      }

      #omitted
      omitted<-!kw%in%mykeywords()
        if (length(omitted)>0){
          text<-paste("omitting the missing keyword(s):",paste(kw[omitted],collapse=", "))
          omit_text(text)
        }

    })

    output$kwwarn = renderText({
      print(omit_text())
         })

    output$kwPlot = renderPlot({
      par(mfrow=c(1,2))
      if(is.igraph(netk())){
      #plot
      coords<-layout_in_circle(netk(), order = V(netk()))
      edge.start <- ends(netk(), es=E(netk()), names=F)[,1]
      edge.col <- V(netk())$color[edge.start]
      plot(netk(), ylim=c(-1,1), xlim=c(-1,1),layout=coords,edge.color=edge.col, edge.curved=0.1, main="Keyword relations")
        legend("bottomleft", legend=paste(1:4,colnames(topic_prop)[1:4]), border=F,pch=21,pt.bg=rev(cols)[1:4],bty="n",cex=0.8)
        legend("bottomright", legend=paste(5:K,colnames(topic_prop)[5:K]), border=F,pch=21,pt.bg=rev(cols)[5:K],bty="n",cex=0.8)
      #show papers
      plot(c(0,1),c(0,1), axes=F, col=NA, main="Top papers", xlab="", ylab="")
      par(xpd=T)
      for(p in 1:length(papers())){
        text(0,1-(p-1)*length(papers())/220, papers()[p], col=pcols()[p],cex=0.8, pos=4)
        }
      }
  }, width="auto")

  },
options=list(height=800)
)
```